---
title: "Tutorial 1"
author: "your name here"
date: "current date here"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Tutorial 1

This assignment is due August 27th. In order to receive full credit students should:

1.  Follow Tutorial 1 from start to finish.
2.  Complete the Questions as they go.
3.  Type their answers into the provided code chunks below or in the text space provided.
4.  Ask questions if they arise.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

**1a. Get and set your working directory based on the suggested file path in Tutorial 0.**

**1b. Install the Palmer penguins data set onto your computer.**

**1c. Import the tidepool metadata practice data set.**

```{r question_1, echo = FALSE}

```

## Question 2

**2a. Retrieve the summary, head, and tail of the tide pool metadata set**.

```{r Question_2}

```

**2b. What's the mean tidepool volume?**

Answer:

**2c. How many rows does the tidepool metadata set have?**

Answer:
