---
title: "R-Preparedness"
author: "your name here"
date: "current date here"
output: html_document
editor_options: 
  chunk_output_type: console
---

# R Preparedness

This assignment is due August 27th. In order to receive full credit students should:

1.  Follow Tutorial 1 from start to finish.
2.  Complete the Questions as they go.
3.  Type their answers into the provided code chunks below or in the text space provided.
4.  Ask questions if they arise.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

**1a. Get and set your working directory based on the suggested file path in Tutorial 0.**

**1b. Install the Palmer penguins data set onto your computer.**

**1c. Import the tidepool metadata practice data set.**

```{r question_1, echo = FALSE}

```

## Question 2

**2a. Retrieve the summary, head, and tail of the tide pool metadata set**.

```{r Question_2}

```

**2b. What's the mean tidepool volume?**

Answer:

**2c. How many rows does the tidepool metadata set have?**

Answer:

## Question 3

**3a. What is the atomic data type of tide pool size? How many tide pool sizes are there?**

```{r Question_3a}

```

**3b. Rename the tidepool metadata object (if you haven't already) using** `snake_case` **convention.**

```{r Question_3b}

```

**3c. Rename all variables in your new tide pool data frame to match** `snake_case` **conventions.**

```{r Question_3c}

```

## Question 4

**Use code to install the tidyverse and dplyr packages; Load the packages into your working session.**

```{r Question_4}

```

*\*\*Note:* This would normally be done at the beginning of any real working script. Ensure that all package installations take place before any commands are run.For the sake of our learning here, we're installing it later into our script.

## Question 5

**5a. Filter the tide pool data to only include large tide pools with a macro_richness that is \>= to 10.**

```{r Question_5a}

```

**5b. Filter the Palmer penguins data set to only include male penguins that were measured in every year except 2007 on Biscoe island (HINT: try using the ! logical operator to exclude 2007).**

```{r Question_5b}

```

## Question 6

**6a. How could you use `arrange()` to sort all missing values to the start? Write a line of code that sorts the missing values (NAs) in the penguins data set to the start. (HINT: try using `is.na()`).**

```{r Question_6a}

```

**6b. In the tide pool data frame, arrange the observations in descending order by the Macro-organism Shannon mean and the Macro-organism richness.**

```{r Question_6b}

```

## Question 7

**7a. Create a sub-set of the tidepool data that only includes the pool number, Shannon's diversity metrics for macro sessile and mobile organisms, and the macro richness.**

```{r Question_7a}

```

**7b. Arrange your new sub-set in descending order by the pool number.**

```{r Question_7b}

```

## Question 8

**8a. Re-import the tidepool metadata set.**

```{r Question_8a}

```

**What effect does this have to the work we've already done to the data set?**

Answer:

**8b. Re-name the tidepool data frame object using `snake_case` once again (if you didn't already do so during re-import). Using dplyr and the `rename()` command, rename all the columns in the tidepool metadata using `snake_case`.**

```{r Question_8b}

```

## Question 9

**Using the tidepool metadata sub-set you created in question 7, add together the Shannon's Diversity for both sessile and mobile macroscopic organisms. Name this new variable appropriately.**

```{r Question_9}

```

## Question 10

**Only using pipes**

**10a. Using the tidepool metadata, group by tide pool size and get the mean Shannon's diversity of both the mobile and sessile macro-organisms. Save this in a new data frame object (a tibble) and name it appropriately.**

```{r Question_10a}

```

**10b. Using the penguins data, create a new sub-set that only includes Chinstrap penguins and their bill measurements. Get the mean of the bill measurements in each year the penguins were measured.**

```{r Question_10b}

```

## Question 11

**11a. Create a scatter plot using the tidepool metadata exploring the macro richness and tidepool volume.**

```{r Question_11a}

```

**11b. Create a bar plot showing the macro richness by tidepool size. (HINT: you will need to use the argument** `stat = 'identity'` **in your** `geom_` **call).**

```{r Question_11b}

```

**11c. Use the** `?geom_bar` **help command to pull up the help page for the** `geom_bar()` **command. Spend a little bit of time reading. You can also look at** `?geom_col`.

```{r Question_11c}
?geom_bar
?geom_col
```

**Why do we need to use the** `stat = 'identity'` **argument when running** `geom_bar()`?

Answer:

**11d. What happens if we don't add stat = 'identity'?**

```{r Question_11d}
# Optional to add code here if you'd like to support your answer
```

Answer:

## Question 12

**Using the tidepool metadata, create a scatter plot that shows the relationship between the sessile and mobile macro-organisms Shannon diversity colored by tidepool size.**

```{r Question_12}

```

## Question 13

**Using the penguins data set, create a scatter plot faceted by species and island that explores the relationship between bill depth and bill length. Add additional aesthetic mappings as wanted or needed.**

```{r Question_13}

```

## Question 14

**Using the penguins data set and a pipe string, create a scatter plot faceted by species and sex that explores the relationship between bill depth and body mass only in the year 2007.**

```{r Question_14}

```

## Question 15

**Using the tidepool metadata, explore the relationship between macro organism richness and tide pool volume between the two tide pool sizes (large and small). Fit a linear regression line to visualize the linear relationship between y \~ x.**

```{r Question_15}

```

*(HINT: This one is tricky and will involve you using the mathematical operator log() on one of the axis. You can do so without piping using dplyr commands OR altering the data frame itself).*
